---
title: "Crime Analysis 2020"
author: "Kahlil Wehmeyer"
date: "11/4/2020"
output:
  pdf_document:
    toc: yes
    latex_engine: xelatex
  html_document:
    toc: yes
    toc_float: yes
subtitle: "Gainesville Office of Strategic Initiatives"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
setwd("~/Documents/gainsville_exercise")
library(tidyverse); theme_set(theme_minimal())
library(knitr)
library(patchwork)
library(janitor)
library(maps)
library(ggridges)
library(lubridate)
library(sf)
```


# Introduction

The goal of this report is to identify crime trends in the Gainesville. 
The hope is to aide in coordinating police more effectively, and generating ideas for how
the community can be enhanced through programs, infrastructure changes and other social measures.


# Data
The data for this exploration is based on the public _"Crime Responses"_ data set which is available [here](https://data.cityofgainesville.org/Public-Safety/Crime-Responses/gvua-xt9q). The version used in this report was last updated on November 3rd, 2020.

Additional information about districts is available [here](https://data.cityofgainesville.org/Geospatial-Maps-/Color-Coded-Gainesville-Police-Zones/qzci-z4wx)


```{r importing data, echo=FALSE, include=FALSE}
if (!file.exists("data/cleaned_responses.csv")) {
  source("cleaning_script.r")
} else {
  responses <- read_csv("data/cleaned_responses.csv")
  areas <- read_sf("data/Gainesville Police Zones.geojson") 
  st_as_sf(responses, coords = c("longitude", "latitude"), crs = st_crs(areas), remove=FALSE) -> responses
}

responses$offense_day_of_week <- factor(responses$offense_day_of_week, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday","Sunday"))
responses$report_day_of_week <- factor(responses$report_day_of_week, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday","Sunday"))
```


```{r district_layouts}
ggplot(data = areas) +
  geom_sf(aes(fill=district))+
  
  labs(title = "Districts") +
  guides(fill=FALSE) +
  theme_void() -> police_districts
ggplot(data = areas) +
  geom_sf(aes(fill=district)) +
  geom_sf(aes(fill = sector)) +
  
  labs(title = "Sectors") +
  guides(fill=FALSE) +
  theme_void() -> police_sectors
ggplot(data = unique(areas)) +
  geom_sf(aes(fill=objectid)) +
  
  # Axis and titles
  labs(title = "Zones") +
  guides(fill=FALSE) +
  theme_void() -> police_zones

(police_districts + police_sectors) + police_zones + 
  plot_annotation(title = "Police",
                  theme = theme(plot.title = element_text(size=30,hjust = 0.5, vjust = 1))) 
```


```{r data_outside_districts}
responses %>% 
  count(in_district) %>% 
  mutate(pct = prop.table(n)) %>% 
  
  ggplot(aes(x = in_district, y = pct, label = scales::percent(pct), fill = in_district)) +
  geom_col() +
  geom_text(position = position_dodge(width = 0.1), vjust = -0.5, size = 4) +
  
  # Labels and Axis
  labs(title = "Percentage of Incidents That Occured Within Police Zones") +
  xlab("In Police Zones") +
  ylab("Percentage") +
  
  guides(fill = FALSE)
```

There is a small percentage of crimes that have locations outside of the given geographic areas. 
A small percentage of the crimes take place in other parts of Florida, with an even smaller
percentage happening out of state.
Our focus is on improving life for the immediate community, therefore those records will be discarded for the time being.

```{r filtering_out_of_district}
responses <- responses %>% filter(!is.na(label))
```





```{r functions}
mcc_zone <- function(zone) {
  
  responses %>% 
    filter(label == zone) %>% 
    group_by(year = year(offense_date), incident_type) %>% 
    summarise(count = n()) %>% 
    group_by(year) %>% 
    top_n(n = 3, wt = count) %>% 
    
    ggplot() +
    geom_col(aes(x = 0, y = count, fill = incident_type), position = "dodge") +
    facet_wrap(~year) +
    
    labs(title = paste0("Most Common Crimes | Zone ",zone)) +
    labs(fill = "Incident Type") +
    
    theme(
      axis.title.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.x = element_blank()
    )
}

area_plot <- function(zone, it = "Burglary to Conveyance"){
  zone_area <- areas %>% filter(label == zone)
  
  zone_reponses <- responses %>% 
    filter(label == zone & incident_type == it) %>% 
    group_by(longitude, latitude) %>% 
    summarise(n = n(), hour = median(as.integer(offense_hour_of_day)))
  
  ggplot() +
    geom_sf(data = zone_area, fill = "darkgray") +
    geom_sf(data = zone_reponses, aes(size = n, color = hour), alpha = 0.7) +
    
    theme(axis.text.x = element_text(angle = 45)) +
    
    labs(title = paste0("Hotspots for Zone ", zone)) +
    labs(subtitle = it) +
    labs(color = "Hour of Day") +
    labs(size = "Count") +
    scale_color_viridis_c(trans = 'reverse')
  
}

time_of_crime <- function(zone, it = "Burglary to Conveyance"){
  
  responses %>% 
    filter(label == zone & incident_type == it) %>% 
    mutate(offense_hour_of_day = as.integer(offense_hour_of_day)) %>% 
    
    ggplot(aes(x = offense_hour_of_day)) +
    geom_density() +
    
    xlab("Hour of Day") +
    ylab("Density") +
    labs(title = paste0("Distribution of Occurences in Zone ", zone)) +
    labs(subtitle = it)
}
day_of_crime <- function(zone, it = "Burglary to Conveyance"){
  responses %>% 
    filter(label == zone) %>% 
    filter(incident_type == it) %>% 
    
    ggplot() +
    geom_histogram(aes(x = offense_day_of_week), stat = "count", fill = "orangered", alpha = 0.8) +
    
    labs(title = "Histogram of Offense Day") +
    labs(subtitle = paste0("Zone ", zone," | ", it)) +
    xlab("Day of Week") +
    ylab("Count")
}
```

# Common Crimes 
Here we will look at a general trend in crimes for all zones **excluding** zone Echo.


We see a general trend in that all the zones have been greatly successful in reducing the number of crimes committed in recent years. Whatever strategy has been implemented in helping reduce crime has been highly effective. However, the crime that seems to prevail in concerning numbers is Burglary to a Conveyance. 

## Time Trends

```{r overall_offense_count}
responses %>% filter(year(offense_date) > 2014) %>% 
  group_by(date = date(offense_date)) %>% 
  summarise(offenses = n()) %>% 
  
  ggplot() + 
  geom_smooth(aes(x = date, y = offenses)) +
  
  labs(title = "Offenses by Date") +
  labs(subtitle = "2015 - Present") +
  xlab("Year") +
  ylab("Number of offenses")
```

This graph displays that the steady crime decrease from 2015 to 2020.

```{r most_common_offenses}
responses %>% 
  filter(year(offense_date) > 2014) %>% 
  group_by(year = year(offense_date), incident_type) %>% 
  summarise(count = n()) %>% 
  group_by(year) %>% 
  top_n(n = 3, wt = count) %>% 
  
  ggplot(aes(x = year, y = count, fill = incident_type)) +
  geom_bar(position = "dodge", stat="identity") +
  
  # Labels
  labs(title = "Top 3 Incidents by Year") + 
  labs(subtitle = "2015 - 2020") +
  xlab("Year") + 
  ylab("Count") +
  labs(fill = "Incident Type") 
  
```

Over the last 5 years the biggest plights to the community has been conveyance burglaries, retail theft, and _other_ forms of petit theft. In agreement with the previous graph, it is evident that these crimes have been reduced in recent years.

```{r}
top_crimes <- responses %>% 
  group_by(incident_type) %>% 
  summarise(n = n()) %>% 
  top_n(12, wt = n)

responses %>% 
  group_by(month = month(offense_date, label = TRUE), incident_type) %>% 
  summarise(n = n()) %>% 
  filter(incident_type %in% top_crimes$incident_type) %>% 
  
  
  ggplot() +
  geom_tile(aes(y = incident_type, x = month, fill = n )) +
  scale_fill_viridis_c() +
  
  labs(title = "Heatmap of Most Prevalent Crimes") +
  xlab("Month") +
  ylab("") +
  guides(fill = guide_legend(title="Count"))
  
```

Burglary to conveyance peaks drastically in July. Retail theft also seems to peak quite drastically in January. All other crimes tend to follow weak patterns throughout the year.


## Burglary to Conveyance

The majority of zones suffer from burglary to conveyance as their most common crime.

The only districts that deviate from this trend are Echo and to a lesser extent Juliet.

```{r}
zone_areas <- areas %>% 
  filter(label != "Echo" & label != "Juliet") 

zone_response <- responses %>% 
  filter(label != "Echo" & label != "Juliet" & incident_type == "Burglary to Conveyance") %>% 
  group_by(longitude, latitude) %>% 
  summarise(n = n(), hour = median(as.integer(offense_hour_of_day))) 

ggplot() +
  geom_sf(data = areas, color = "darkred", fill = "darkgray") +
  geom_sf(data = zone_areas) +
  geom_sf(data = zone_response, aes(color = hour), alpha = 0.2) +
  
  labs(title = "Heatmap of Crimes") +
  labs(subtitle = "Burglary of Conveyance | Excluding Echo and Juliet") +
  scale_color_viridis_c(trans = "reverse") +
  
  guides(color = guide_legend(title="Hour of Day", override.aes = list(alpha = c(1), size = c(5))))

```

Observe that the majority of points on the map are dark, indicating that the crime occurred at a later hour in the day. Note the distribution of when these crimes are committed for all zones.

```{r street_lights}
responses %>% 
  filter(incident_type == "Burglary to Conveyance") %>% 
  mutate(offense_hour = as.integer(offense_hour_of_day)) %>% 
  
  ggplot(aes(x = offense_hour)) +
  geom_density() +
  geom_histogram(alpha = 0.4, fill = "gold", aes(y=..density..)) +
  
  # Labels and Titles
  labs(title = "When do burglary of conveyances happen?") +
  labs(subtitle = "For all zones") +
  xlab("Hour of Day") +
  ylab("Density") +
  
  # Annotations
  annotate("rect",
           ymin = 0.04, ymax = 0.1,
           xmin = 17, xmax = 22,
           alpha = 0.2) +
  
  annotate("text", label = "Darker hours",
           x =  7.5, y = 0.06) +
  
  annotate("curve",
           x = 10, xend = 15.5,
           y = 0.06, yend = 0.07,
           curvature = -0.2,
           arrow = arrow(length = unit(2,"mm"))) +
  
  annotate("curve",
           x = 5, xend = 1,
           y = 0.06, yend = 0.07,
           curvature = 0.2,
           arrow = arrow(length = unit(2,"mm"))) 

```

It is possible that if we provide better street side illumination, not only would it deter conveyance burglaries, it would also enhance the safety of pedestrians, increase civilian surveillance as well as increase driver awareness in streets, parking lots, and driveways.

## Petit Theft - Retail

```{r}
zone_areas <- areas %>% 
  filter(label == "Echo" | label == "Juliet") 

zone_response <- responses %>% 
  filter(label == "Echo" | label == "Juliet" & incident_type == "Theft Petit - Retail") %>% 
  group_by(longitude, latitude) %>% 
  summarise(n = n(), hour = median(as.integer(offense_hour_of_day))) 

ggplot() +
  geom_sf(data = areas, color = "darkred", fill = "darkgray") +
  geom_sf(data = zone_areas) +
  geom_sf(data = zone_response, aes(color = hour), alpha = 0.25) +
  
  labs(title = "Heatmap of Crimes") +
  labs(subtitle = "Theft Petit - Retail | Excluding Echo and Juliet") +
  scale_color_viridis_c(trans = "reverse") +
  ylim(29.58, 29.68) +
  
  
  guides(color = guide_legend(title="Hour of Day", override.aes = list(alpha = c(1), size = c(5))))
```

Zone Echo (bottom left) accounts for 30% of all petit retail theft. In the drill down later it's observed that there are 3 stores that almost account entirely for the retail theft. They are: 

- Walmart
- Lowe's
- Target


# Deep Dive into Zones

Taking a deeper dive, dissecting the crime patterns zone by zone will give a better idea of how we might help to reduce crime and increase community safety.

## Alpha

```{r}
mcc_zone("Alpha")
```

Since 2015 the rate of conveyance burglaries has slowly declined, however it still continues to be the main source of crime in zone Alpha.

```{r}
area_plot("Alpha") + 
  annotate("segment", color = "orangered", size = 1.1,
           x = -82.38, xend = -82.38,
           y = 29.673, yend = 29.71) +
  annotate("segment", color = "orangered", size = 1.1,
           x = -82.38, xend = -82.36,
           y = 29.71, yend = 29.715) + 
  annotate("rect", color = "orangered", alpha = 0.1,
           xmin =  -82.405, xmax = -82.386,
           ymin = 29.718, ymax = 29.732)
```

The orange-ish lines indicate suggested police routes for this graph and the subsequent graphs.
The rectangle indicates a hotspot of conveyance crimes. This would be useful for the police responsible for patrolling this zone by guiding their routes to optimize their coverage.

```{r}
day_of_crime("Alpha") / time_of_crime("Alpha")
```

The distribution of time that the burglaries occur agrees with the general trend shown earlier in the report; darker hours of the day lead to more crimes. 
There isn't a significantly prominent day of the week that the crimes tend to occur.

## Bravo

```{r}
mcc_zone("Bravo")
```

Bravo does not show a strong trend in crime change.
However, observe that in recent years, burglary to conveyance has remained prominent and has even seen a slight increase from 2019.

```{r}
area_plot("Bravo") + 
  annotate("segment", color = "orangered",
           x = -82.368, xend = -82.34,
           y = 29.702, yend = 29.668)
```

This heatmap shows a somewhat even distribution of crime that is skewed towards the North. A decent strategy could be to position police officers along the NW/SE diagonal from 29.705°N to 82.34°W. This way the most of the area is covered, or is within relatively close distance.
Note these orange lines are created by hand are not exactly drawn to reduce residual distance from crime centers.



```{r}
day_of_crime("Bravo") / time_of_crime("Bravo")
```

Sharing similarities with zone Alpha, and the overall trend, darker hours
lead to more crimes
Additionally, observe a general trend that burglaries tend to occur later in the week. However,
this is a very weak correlation.

## Charlie
```{r}
mcc_zone("Charlie")
```

Zone Charlie has seen a massive decline in crime since 2018.
There seems to be a nearly equal share of both burglaries to 
conveyance, criminal mischief, and damage to property in 2020.
Clearly whatever program or strategy was implemented to reduce burglary
to conveyance in Zone Charlie should be considered, if possible,
in other zones suffering from similar plights.

### Damage to property
```{r}
area_plot("Charlie", "Damage to Property") +
  annotate("segment", color = "orangered",
           x = -82.399, xend = - 82.373,
           y = 29.659, yend = 29.651) +
  
  annotate("rect", color = "orangered", alpha = 0.2,
           xmin = -82.35, xmax = -82.339,
           ymin = 29.652, ymax = 29.666)
```

The majority of cases for damage to property occur in the Southeast corner of the area.
Positioning a considerable amount of officers here would be effective.
Having a few officers occasionally patrolling along the NW/SE diagonal
from 29.66°N to 82.375°W would be effective.
Seeing as this diagonal could be covered by just a few officers and thwart those remaining cases of damage to property.

```{r}
day_of_crime("Charlie", "Damage to Property") / time_of_crime("Charlie", "Damage to Property")
```

Note that there are two high points in the day for damage to property. For a short period in the early morning hours, then again around mid-day, from 12-5PM, tapering off as the night goes on.
Again, it would be recommended that officers patrol the designated hot spots primarily during these hours to ensure the most effective use of their time.

It seems that Friday is the somewhat most prominent day for these crimes to occur. This is likely
due to the fact that it's the start of the weekend. Children don't have school and people
are drinking.

### Criminal Mischief

```{r}
area_plot("Charlie", "Criminal Mischief (misdemeanor)") +
  annotate("segment", color = "orangered",
           x = -82.399, xend = - 82.373,
           y = 29.659, yend = 29.651) +
  
  annotate("rect", color = "orangered", alpha = 0.2,
           xmin = -82.35, xmax = -82.339,
           ymin = 29.652, ymax = 29.666)
```

Similarly to the damage to property map, it is recommend that officers focus their attention to the Southeast corner of this zone with a few officers doing routine patrolling around the remainder of the zone.

```{r}
day_of_crime("Charlie", "Criminal Mischief (misdemeanor)") / time_of_crime("Charlie", "Criminal Mischief (misdemeanor)")
```
 
Criminal mischief follows similar time trends as damage to property. Following roughly the same schedule as the previous suggestion, officers should be able to cover those areas effectively.

 
 
 ## Burglary of Conveyance
```{r}
area_plot("Charlie") +
  annotate("segment", color = "orangered",
           x = -82.399, xend = - 82.373,
           y = 29.659, yend = 29.651) +
  
  annotate("rect", color = "orangered", alpha = 0.2,
           xmin = -82.35, xmax = -82.339,
           ymin = 29.652, ymax = 29.666)
```
 
Yet again, a fair concentration of officers in the Southeast corner is advisable. Additionally, having a number of officers patrolling the NW/SE diagonal from 29.675°N to 82.38°W would cover the majority of instances. It's also advisable that officers patrolling that diagonal skew their routes south to maintain proximity to the central cluster of crimes.

```{r}
day_of_crime("Charlie") / time_of_crime("Charlie")
```

Continuing the trend of crimes being committed later in the day, starting around dusk.

## Delta
```{r}
mcc_zone("Delta")
```

Retail theft has greatly been reduced in recent years with the new common culprit being burglary to conveyance for zone Delta. Emphasis will be placed on this crime since it is most prevalent.

```{r}
area_plot("Delta") +
  annotate("segment",
           x = -82.413, xend = -82.39,
           y = 29.66, yend = 29.635, color = "orangered") +
  
  annotate("segment",
           x = -82.39, xend = -82.375,
           y = 29.635, yend = 29.635, color = "orangered")
```

Due to the somewhat awkward nature of this area, it would be advised that officers patrol the central line of the polygon starting from 29.66°N, 82.41°W moving southeast to 29.635°N, 82.39°W and then eastbound along 29.634°N.
The over-arching trend seems to be that the burglaries tend to be focused near the southeast quadrant.

```{r}
day_of_crime("Delta") / time_of_crime("Delta")
```

The trend continues to follow that of the previous zones and the overall trend. Crimes happen 
at night time.
It's also interesting to note that the day of the week graph skews slightly to Sunday.
However, the difference is less than 50 incidents from Mon-Wed so it is not recommended to make any
massive scheduling decisions based on this information.

## Echo
```{r}
mcc_zone("Echo")
```

Breaking from the trend of the primary crime being burglary to conveyance, zone Echo has a massive problem with retail petit theft. Zone Echo alone accounts for nearly 30% of all petit theft among all zones.
The crime rate has decreased slightly since 2019 but clearly it has been difficult to drastically subdue the number of these crimes.

```{r}
area_plot("Echo", "Theft Petit - Retail")
```

The heatmap does not draw a representative picture of the data. There is large share of retail crimes in this zone, yet if compared it to some of the other maps it looks like this is possibly the safest zone. However, observing the legend for crime count on the left leads warrants examination of those locations more in depth.

```{r}
responses %>% 
  filter(label == "Echo" & incident_type == "Theft Petit - Retail") %>% 
  group_by(address) %>% 
  summarise(n =  n()) %>% 
  top_n(n = 10, wt = n) %>% 
  
  ggplot() +
  geom_col(aes(x = address, y = n), fill = "orangered") + 
  coord_flip() +
  
  labs(title = "Theft Counts by Address") +
  labs(subtitle = "Zone Echo") +
  xlab("Address") +
  ylab("No. of Thefts") +
  
  annotate("text", label = "Walmart",
           y = 750, x = 1, color = "white", size = 5.5) +
  
  annotate("text", label = "Lowe's",
           y = 450, x = 6, color = "white", size = 5.5) +
  
  annotate("text", label = "Target",
           y = 225, x = 10, color = "white", size = 5.5) +
  
  annotate("text", label = "This address points to a Pollo Tropical.\n However it's likely that the crimes actually\n occured at a nearby Target.",
           y = 1000, x = 9) +
  
  annotate("curve", arrow = arrow(length = unit(2,"mm")),
           y = 550, yend = 430,
           x = 10, xend = 10,
           curvature = 0.2)
```

The majority of thefts occur at Walmart, Lowe's and Target. All of these are within relatively close proximity in the shopping district. The question is, how can theft be prevented in stores? 

```{r}
day_of_crime("Echo", "Theft Petit - Retail") / time_of_crime("Echo", "Theft Petit - Retail")
```

The offenses for retail theft are concentrated around 3PM. There is no clear pattern 
in the day of the week in which these crimes occur.

## Foxtrot

```{r}
mcc_zone("Foxtrot")
```

Foxtrot observes a fairly steady level of crime for burglaries to conveyance. It is worth noting 
that this area has not seen a drastic decrease in crime, as is observed in other zones.


```{r}
area_plot("Foxtrot") + 
  annotate("segment", color = "orangered", size = 1.1,
           x = -82.37, xend = -82.36,
           y = 29.628, yend = 29.633) +
  
  annotate("rect", color = "orangered", alpha = 0.3,
           xmin = -82.37, xmax = -82.355,
           ymin = 29.613, ymax = 29.623)
```

The hostpots are concentrated in the center with a small belt of crime on the Northern wall.    

```{r}
day_of_crime("Foxtrot") / time_of_crime("Foxtrot")
```

There is a continuation of the trend that burglaries to conveyance happen primarily during
the darker hours of the day. 
There is weak skew on the day of the week histogram, with more cases are observed in the beginning of the week.

## Golf

```{r}
mcc_zone("Golf")
```

There has been a drastic theme shift in crime for Golf over the last 6 years.
There has been a change from trespassing being a main crime to conveyance burglary being the most prevalent crime.
Additionally, domestic battery has prevalent in this zone for 2020; historically as well.

### Burglary to Conveyance

```{r}
area_plot("Golf") +
  annotate("rect", color = "orangered", alpha = 0.3,
           xmin = -82.34, xmax = -82.314,
           ymin = 29.665, ymax = 29.688) 
```

The area of focus should be the lower sections of this zone. There is also a small hotspot near the eastern enclave.    

```{r}
day_of_crime("Golf") / time_of_crime("Golf")
```

Unsurprisingly the theme for burglary of conveyance during later hours is continued.
No particular day is correlated with crime.

### Domestic Simple Battery

```{r}
area_plot("Golf", "Domestic Simple Battery") +
  annotate("rect", color = "orangered", alpha = 0.3,
           xmin = -82.33, xmax = -82.318,
           ymin = 29.684, ymax = 29.695) +
  annotate("curve", color = "orangered", size = 1,
           x = -82.337, xend = -82.32,
           y = 29.684, yend = 29.675,
           curvature = 0.8)
```

There is a notable hotspot in the center, signified by the orange rectangle, and a small belt that is
is centrally outlined by the curve.

```{r}
day_of_crime("Golf", "Domestic Simple Battery") / time_of_crime("Golf", "Domestic Simple Battery")
```

There isn't a particularly strong pattern in the time that these crimes occurs.
Generally, domestic battery incidents happened later in the evening and there is not
strong relationship with the day of the week.


## Hotel

```{r}
mcc_zone("Hotel")
```

An interesting trend in zone Hotel is that Dcf investigations have been a major
issue for a number of years. Additionally, the common crime of conveyance burglary shows up
as a notable issue.

### Dcf Investigations

```{r}
area_plot("Hotel", "Dcf Investigations") +
  annotate("rect", color = "orangered", alpha = 0.3,
           xmin = -82.315, xmax = -82.304,
           ymin = 29.667, ymax = 29.688) +
  
  annotate("rect", color = "orangered", alpha = 0.3,
           xmin = -82.28, xmax = -82.274,
           ymin = 29.677, ymax = 29.682) +
  
  annotate("rect", color = "orangered", alpha = 0.3,
           xmin = -82.264, xmax = -82.258,
           ymin = 29.673, ymax = 29.68)
```

There are three major hotspots for dcf investigations. Since these investigation are 
social issues, it is important to consider what can be done to help families care for their children.
It would be difficult to justify an increase in police patrolling as a means to improve
the quality of life for children in troubled homes.
Therefore it is recommended that the community and local government help put in place programs
to improve the overall environment in which these children are raised.

Since Dcf investigations aren't necessarily a "on the spot" crime, much insight would not be gained by 
observing the time in which these investigations happen.

### Burglary to Conveyance 

```{r}
area_plot("Hotel") +
  annotate("rect", color = "orangered", alpha = 0.3,
           xmin = -82.315, xmax = -82.301,
           ymin = 29.667, ymax = 29.693) 
```

The majority of crimes occur in the western region of the zone. There are a couple small spots that are nearby that also indicate issues, but the brunt of the crime happens within the rectangle.

```{r}
day_of_crime("Hotel") / time_of_crime("Hotel")
```

Yet again, observe that conveyance burglaries occur later in the day and hold
no strong relationship with the day of the week histogram.

## India

```{r}
mcc_zone("India")
```

India has seen a massive decrease in crime since 2015. The main crimes were trespassing and retail theft. In 2020, for the most part, those have been reduced greatly. Now the main issue is 
burglary to conveyance.

```{r}
area_plot("India")
```

The overall spread is distributed fairly evenly across
this area. Even the largest "hotspots" only account for 20 burglaries.
Due to the sparsity of these crimes it could be surmised that increased police surveillance would 
not make a drastic difference.

```{r}
day_of_crime("India") / time_of_crime("India")
```

The assumptions hold true to the overall trend,
these crimes happen later in the day and it would be
recommend to improved streetside lighting.

## Juliet

```{r}
mcc_zone("Juliet")
```

As stated in the earlier sections of the report. Juliet sees a high number of
petit retail theft. The numbers of these thefts have been reduced considerably in the 
last year, even though it continues to be the main issue.


```{r}
area_plot("Juliet", "Theft Petit - Retail")
```

Similar to Echo, the hotspots seem to indicate a few key locations where the theft occurs.

```{r}

responses %>% 
  filter(label == "Juliet" & incident_type == "Theft Petit - Retail") %>% 
  group_by(address) %>% 
  summarise(n =  n()) %>% 
  top_n(n = 10, wt = n) %>% 
  
  ggplot() +
  geom_col(aes(x = address, y = n), fill = "orangered") + 
  coord_flip() +
  
  labs(title = "Theft Counts by Address") +
  labs(subtitle = "Zone Echo") +
  xlab("Address") +
  ylab("No. of Thefts") +
  
  annotate("text", label = "Walmart", color = "white", size = 5,
           y = 1050, x = 6)
```

The skew here is quite jarring. Walmart accounts almost entirely for all retail thefts.
Similarly zone Echo, it is unsure if there is a simple solution to preventing in store crimes.

```{r}
day_of_crime("Juliet",  "Theft Petit - Retail") / time_of_crime("Juliet", "Theft Petit - Retail")
```

The crimes happen typically later in the day, and closer to the weekend.

## Mike 

```{r}
mcc_zone("Mike")
```

Zone Mike has also seen a drastic decrease in crime over the last couple years. Currently the biggest issue is, oddly, bicycle theft.

```{r}
area_plot("Mike", "Theft Petit - Bicycle") +
  annotate("segment", color = "orangered",
           x = -82.3356, xend = -82.3356,
           y = 29.6428, yend = 29.654)
```

See that the spread of the crimes is fairly even with a majority of the crimes happening 
in proximity to orange line.

```{r}
day_of_crime("Mike", "Theft Petit - Bicycle") / time_of_crime("Mike", "Theft Petit - Bicycle")
```

Bicycle thefts peak in the early hours of the morning with no strong relationship to the day of the 
week.

## November 

```{r}
mcc_zone("November")
```

For our final zone there is a steady decrease in crimes with the most common issue being
conveyance burglary and domestic battery. However, the volume of these crimes is relatively low for both, which does not prompt further analysis.


# Conclusions

Although crime has been greatly reduced in recent years there is still a big issue with burglary to conveyance and petit theft. For burglary to conveyance it is recommended that enhanced street side illumination be developed in the problem areas that have been identified. The benefits are numerous,
extending past just crime prevention.

For petit theft crimes it seems that there are 4 retail establishments that are the main areas where
these crimes occur. Reducing retail theft is a complicated issue that doesn't have any simple solutions
involving coordinating police within stores or around shopping districts.
A good start may be to develop the local economy further and develop programs to distance the community from impoverishment.
It should be acknowledged that looting and rioting due to 2020 political unrest likely has also contributed heavily to retail theft numbers.

Furthermore, the massive spike in unemployment also gives further reason to the observation of higher retail theft.

Dcf investigations are also a concerning problem in Zone Golf. Community programs should be explored 
with the expressed goal of enabling parents to take better care of their children, and allowing  
children to have better resources and safety if they belong to troubled homes.


## Further Opportunities

* Use street data to plot out police routed more precisely.
* Research into causes and methods of petit theft prevention.
* Forecasting crime zones and areas for future planning.
* Use COVID-19 data in concert with crime data to possibly predict and model changes in criminal behavior.

```{r eval=FALSE}

responses %>% 
  filter(year(offense_date) == 2020) %>% 
  filter(incident_type == "Burglary to Conveyance" | incident_type == "Theft Petit - Retail" |
           incident_type ==  "Dcf Investigations") %>%
  group_by(month = month(offense_date, label = TRUE), incident_type)  %>% 
  summarise(count = n()) %>% 
  
  ggplot() +
  geom_col(aes(x = month, y = count, fill = incident_type), position="stack") +
  
  labs(title = "Problematic Crimes for 2020") +
  labs(subtitle = "2020 Only") +
  xlab("Month") +
  ylab("Count") +
  guides(fill = guide_legend(title = "Incident Type"))

```



