---
title: "City Of Gainseville"
author: "Kahlil Wehmeyer - Office of Strategic Initiatives"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Documents/gainsville_exercise")
library(tidyverse); theme_set(theme_minimal())
library(knitr)
library(janitor)
```

# Introduction

# Data
The data for this exploration is based on the public _"Crime Responses"_ data set which is available [here](https://data.cityofgainesville.org/Public-Safety/Crime-Responses/gvua-xt9q). The version used in this report was last updated on November 3rd, 2020.

Additional information about districts is available [here](https://data.cityofgainesville.org/Geospatial-Maps-/Color-Coded-Gainesville-Police-Zones/qzci-z4wx)

**Note**:\n
Data is being prepared by an external script.\n
- Correcting date types
- Better variable naming
- Removing extraneous variables


```{r importing data, echo=FALSE, include=FALSE}
source('cleaning_script.r')
```

```{r initial look}
head(responses) 
summary(responses)
```


